Predictive Quantization for MIMO-OFDM SVD Precoders using Reservoir Computing Framework

### 2019-04-28 paper/main.tex
Add changes

The output of the reservoir is $\bp_{i,t}$, a vectorized representation of $\bP_{i,t}$.
Therefore, the output of the reservoir is also a $D_{\text{in}}$ dimensional vector ($D_{\text{out}}=D_{\text{in}}$).
The output $\bp_{i,t}$ is obtained using a matrix transformation of the reservoir state $\br_{i,t}$ with the $D_{\text{resv}}\times D_{\text{out}}$ output coupler matrix $\bW^{\text{out}}_{i,t-1}$ viz.
\begin{equation}
\bp_{i,t}=\bW^{\text{out}}_{i,t-1}\br_{i,t}
\label{pred_op}
\end{equation}
\begin{figure*}[ht]
\centering
\includegraphics[width=0.8\textwidth]{images/system2.pdf}
\caption{System Overview: The reservoir computing framework takes the precoder estimate at $t-1$ (viz. $\tilde{\bu}_{i,t-1}$) as input, and executes the forward pass of the predictive quantization framework, shaded in green, to obtain $\tilde{\bu}_{i,t}$. Both the transmitter and receiver utilize the same reservoir (initialized identically), with a minute difference indicated by the cyan shaded block. In the cyan shaded block, receiver finds the appropriate codeword for the $q_P$ function by comparing the prediction with the actual $\bU_{i,t}$, and communicates $\text{fb}_{i,t}$ to allow estimation of $\tilde{\bu}_{i,t}$ at the transmitter. Then, both the receiver and transmitter execute the backward pass (shaded in blue) to optimize $\mathcal{O}_{i,t}$ \eqref{oit}, to ensure identical evolution of both the reservoirs}
\label{res_overview}
\vspace{-10pt}
\end{figure*}
This completes the forward prediction to obtain $\bP_{i,t}$ using only the past observed $\bU_{i,t-1}$ (i.e. $p=1$, Section \ref{section2}). That is,
\begin{equation}
\bP_{i,t}=f_P(\tilde{\bU}_{i,t-1})=\bW^{\text{out}}_{i,t-1}\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})\footnote{For brevity, $\bp_{i,t}$ and $\bP_{i,t}$ are treated to be one and the same}
\label{fwd_pred}
\end{equation}
However, a complete discussion of the prediction algorithm also involves a backward pass, involving the updation of output couplers $\bW^{\text{out}}_{i,t}$, which would follow subsequently.
\subsection{Quantization function $q_P$}
\label{qp}
Before proceeding, it is critical to describe the quantization function $q_P(\bP_{i,t},\text{fb}_{i,t})$, which is similar to the one used in \cite{6891198,Gupt1905:Predictive,6545375}.
Central to the quantization algorithm is the fact that tangent spaces local to a point in manifold are linear vector spaces.
Given two points $\bX$ and $\bY$ in $\text{St}(N_T,N_R)$, a lifting operation $\bT^Y_X=\texttt{lift}(\bX,\bY)$, $\texttt{lift}:\text{St}(N_T,N_R)\times \text{St}(N_T,N_R) \to \mathcal{T}_X \text{St}(N_T,N_R)$ yields a tangent from $X$ to $Y$, denoted as $\bT^Y_X \in \mathcal{T}_X \text{St}(N_T,N_R)$, where $\mathcal{T}_X \text{St}(N_T,N_R)$ is the local tangent space at $\bX$.
A corresponding retraction operation $Y=\texttt{retract}(X,\bT^Y_X)$, $\texttt{retract}: \text{St}(N_T,N_R)\times \mathcal{T}_X \text{St}(N_T,N_R) \to \text{St}(N_T,N_R)$ gives back the manifold point obtained by traversing in the tangent direction given by the second argument.
In this work, the chosen lifting-retraction operations are the Cayley Exponentials, discussed in \cite{DBLP:journals/corr/abs-1708-00045,Gupt1905:Predictive}. % and used in \cite{Gupt1905:Predictive}.
The Cayley exponential lifting operation maps two points in $\text{St}(N_T,N_R)$, to a $N_T\times N_T$ skew Hermitian matrix that represents the tangent from first point to the other.

The quantization algorithm exploits the linear vector space property of the tangent space and quantizes the local tangent space at the predicted precoder $\bP_{i,t}$, viz $\mathcal{T}_{\bP_{i,t}} \text{St}(N_T,N_R)$.
The codebook for $\mathcal{T}_{\bP_{i,t}} \text{St}(N_T,N_R)$ corresponds to a collection of codewords that represent the different directions (tangents) in $\mathcal{T}_{\bP_{i,t}} \text{St}(N_T,N_R)$.
This codebook is also referred to as the base codebook ($\text{base}^{\mathcal{C}}$) for the quantization function.
The feedback from the receiver indicates the optimum tangent in the $\text{base}^{\mathcal{C}}$, which the transmitter can choose to get closest (in terms of chordal distance metric, $d_s(\bX,\bY)$ for $\bX,\bY \in \text{St}(N_T,N_R)$ detailed in \cite{Gupt1905:Predictive,6891198}) to the actual value $\bU_{i,t}$.
With the optimum direction chosen, the next step is to determine how much to move in that particular chosen direction, viz. the length of the chosen tangent direction.
For this, we adopt the strategy in \cite{6891198}, which
controls the magnitude of tangent steps, by having two codebooks
$T^{\mathcal{C}}_p,T^{\mathcal{C}}_m$ of different spreads $s_p,s_m$, but same $2^{\mathcal{B}-1}$ base vectors in the $\text{base}^{\mathcal{C}}$, i.e. $T^{\mathcal{C}}_{\{p/m\}}=s_{\{p/m\}}\text{base}^{\mathcal{C}}$ (All codewords $\in \text{base}^{\mathcal{C}}$ multiplied by $s_{\{p/m\}}$ individually).
The two codebooks are concatenated to form a $2^\mathcal{B}$ length codebook $T^\mathcal{C}$.
The receiver finds the optimal index $\text{fb}_{i,t} \in \{1,2,\ldots,2^{\mathcal{B}}\}$ in $T^\mathcal{C}$ by comparing the chordal distance metric ($d_s$), of each codeword to the actual precoder $\bU_{i,t}$ obtained from the SVD of the channel matrix, using \eqref{eqn:cn1}. The receiver then feeds back $\text{fb}_{i,t}$ to the transmitter using $\mathcal{B}$ bits. The transmitter uses the fed back $\text{fb}_{i,t}$ and \eqref{eqn:cn2} to calculate $\tilde{\bU}_{i,t}$,
\begin{equation}
\label{eqn:cn1}
\text{fb}_{i,t} \gets \text{argmin}_{i \in \{1,2,\ldots,2^{\mathcal{B}}\}}\left (d_s\left (\bU_{i,t},\text{retract}(\bP_{i,t},T^\mathcal{C}[i])\right ) \right)
\end{equation}
\begin{equation}
\tilde{\bU}_{i,t}=q_P(\bP_{i,t},\text{fb}_{i,t})=\text{retract}(\bP_{i,t},T^\mathcal{C}[\text{fb}_{i,t}])
\label{eqn:cn2}
\end{equation}

\noindent $\text{fb}_{i,t}$ is also used to update the spread of the codebooks $T^{\mathcal{C}}_p,T^{\mathcal{C}}_m$.
Depending on whether $\text{fb}_{i,t} \geq 2^{\mathcal{B}-1}$,
i.e. whether the optimum codeword is in $T^{\mathcal{C}}_p$ or
$T^{\mathcal{C}}_m$, the scale parameter $s[k]$, which in turn
controls values of $s_p, s_m$, is updated accordingly,

@@

\noindent Having described the technicalities of $f_P,q_P$, we are now ready to describe the training process for the reservoir.
Recall that to obtain $\tilde{\bU}_{i,t}$ from  $\tilde{\bU}_{i,t-1}$, we have,

@@

The key idea here is that as the predicted matrix $\bP_{i,t}$ (the coarse estimate), gets closer to $\tilde{\bU}_{i,t}$ (the refined estimate), the receiver has to quantize even smaller subspaces and can thus provide a more refined estimate from the feedback $\text{fb}_{i,t}$.
Hence, for the time instant $t\geq1$, we train the reservoir output coupler such that the following objective function is optimized,

@@

here $\lambda>1$ is the history parameter. We wish to perform the following optimization in order to find the optimum $\bW_{i,t}^{\text{out}}$ from the available estimates till time $t$,

@@

Computing the gradient $\frac{\partial \mathcal{O}_{i,t}}{\partial \bW_{i,t}^{\text{out}}}$, setting it to null matrix gives,

@@

where $\texttt{outer}(\bx,\by)$ is the outer product between $N_x$ dimensional vector $\bx$ and $N_y$ dimensional vector $\by$ to yield a $N_x\times N_y$ matrix. Let $\sum_{s=1}^{t}\frac{1}{\lambda^{t-s}}\texttt{outer}(\tilde{\bu}_{i,s},\br_{i,s})$ be $\text{\textbf{num}}_{i,t}$ and $\sum_{s=1}^{t}\frac{1}{\lambda^{t-s}}||\br_{i,s}||^2$ be $\text{den}_{i,t}$, with $\bW_{i,t}^{\text{out}}=\frac{\text{\textbf{num}}_{i,t}}{\text{den}_{i,t}}$. Observe that,

@@

\eqref{wout_upd} makes training of the reservoir easy to implement, and of low complexity as well.
We only need to store the matrix $\text{\textbf{num}}_{i,t-1}$, a number $\text{den}_{i,t-1}$, and using the current reservoir state $\br_{i,t}$ and $\tilde{\bu}_{i,t}$ obtained from \eqref{fpqp}, $\bW_{i,t}^{\text{out}}$ can be calculated, to obtain predicted coarse estimate of $\tilde{\bu}_{i,t+1}$.
This completes our technical discussion for the reservoir computing based framework for predictive quantization illustrated in Fig. \ref{res_overview}

@@

To conclude this section, we now summarize and highlight the novel points of the proposed framework for predictive quantization.
The first key advantage of the proposed framework, over the past work \cite{Gupt1905:Predictive,6891198,6545375} is that the prediction function and quantization function are not independent of each other.
To see this, observe \eqref{update_Wt1} to obtain $\bW^{\text{out}}_{i,t-1}$ from $\tilde{\bU}_{i,t-1}=q_P(f_P(\tilde{\bU}_{i,t-2}),\text{fb}_{i,t-1})$, and \eqref{pred} to obtain $\bP_{i,t+1}$,
\begin{equation}
\highlight{\bW_{i,t-1}^{\text{out}}}=\frac{\frac{\text{num}_{i,t-2}}{\lambda}+\texttt{outer}(\highlight{q_P}(f_P(\tilde{\bU}_{i,t-2}),\text{fb}_{i,t-1}),\br_{i,t-1})}{\frac{\text{den}_{i,t-2}}{\lambda}+||\br_{i,t-1}||^2}
\label{update_Wt1}
\end{equation}
\begin{equation}
\highlight{\bP_{i,t}}=f_P(\tilde{\bU}_{i,t-1})=\highlight{\bW^{\text{out}}_{i,t-1}}\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\highlight{\tilde{\bu}_{i,t-1}})
\label{pred}
\end{equation}
The dependence between $f_P$ and $q_P$ in our framework is in the sense that the backward pass ensures that prediction function `learns' the correlations brought about by $q_P$, in addition to temporal correlations. %Better framing possible
This is done by optimizing $\bW_{i,t}^{\text{out}}$ such that the prediction output to be as close to the output from $q_P$ (Objective function $\mathcal{O}_{i,t}$).
The updated $\bW_{i,t}^{\text{out}}$ is then used in forward pass to predict $\bP_{i,t+1}$, which is hopefully closer to $\tilde{\bU}_{i,t+1}$ than $\bP_{i,t}$ was to $\tilde{\bU}_{i,t}$ due to optimization of $\mathcal{O}_{i,t}$.

\textcolor{blue}{Another advantage of the proposed scheme comes from the ease of training the reservoir, entailed by \eqref{res_upd}.
By storing just one matrix $\text{\textbf{num}}_{i,t}$ and a number $\text{den}_{i,t}$, we capture the entire history of the precoding matrices, with past values weighted by $\lambda$.
This is a departure from previous schemes \cite{Gupt1905:Predictive,6891198,6545375} which store the past $n$ (which has to be pre-decided) precoders in a $n$ sized cyclic buffer kind-of arrangement.
Reservoir computing, thus proposes an easier to train (data centric) scheme, which is not typical for typical data centric ML based schemes.}

\section{Simulation Results}
\label{section4}
\subsection{Simulation setting considered}
\label{setting}
\noindent We simulate both our predictive quantization algorithm and the one presented in \cite{6891198} for the IEEE Pedestrian-A channel, with $N_T=4$, $N_R=2$. 
The channel matrices $\bH_{i,t}$ are generated using Jake’s model for realistic simulations, from the IT++ library through the python wrapper py-itpp \cite{ViditPy}. 
Although in both our work and \cite{6891198}, only temporal correlations are exploited, we compare the BER, quantization error and logsum rate results for the complete MIMO-OFDM setting with number of subcarriers $N=64$, and evenly spaced $8$ fed back subcarriers, indexed from 0, viz. $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$). The precoding matrices at non fedback subcarriers ($i\neq\{9k\}$) are estimated via the Cayley Interpolation method presented in \cite{Gupt1905:Predictive}. Feedback is available for $i=\{9k\}$ subcarriers (i.e. transmitter obtains $\text{fb}_{\{9k\},t}$) for predictive quantization. The setting is visually illustrated in Fig. \ref{table}. A 6 bit codebook generated via Lloyd codebook algorithm for the Stiefel Manifold in \cite{6678348} is used to feedback $\tilde{\bU}_{\{9k\},0}$ and differential quantization (i.e. $\bP_{i,t}=\tilde{\bU}_{i,t-1}$) is used for 10 time instances to provide provide the initial training data for the reservoir and for overcoming initial transient response of the reservoir discussed in \cite{mosleh2017brain}. A 6 bit base codebook is used for the quantization function $q_P$, generated according to the method presented in Section \ref{qp} and \cite{Gupt1905:Predictive}. Hence, for both initialization and subsequent feedback, $\text{fb}_{i,t}$ can be encoded via 6 bits (i.e. $\mathcal{B}=6$ Section \ref{section2}). The bit budget per OFDM frame is thus 48 bits, for total 8 equally spaced fed back subcarriers $\{9k, k \in \{0,1,\ldots,7\}\}$.

For predicting each of the $8$ fed-back subcarriers, we consider $8$ separate reservoirs, each of them evolved separately, with $i\in\{9k, k \in \{0,1,\ldots,7\}\}$ in both the forward pass prediction and quantization ($f_P,q_P$) \eqref{fpqp} and backward pass to update $\bW_{i,t}^{\text{out}}$ \eqref{backpass}. Unlike \cite{Gupt1905:Predictive}, we do not exploit frequency correlations for predictive quantization, and the $8$ reservoirs for prediction of $8$ fed-back subcarriers are kept independent of each other. All these 8 reservoirs are initialized with $\bA_{i\in\{9k\}}$ being an Erdos-Reny\"i graph, with probability of edge connections being $0.2$, $\lambda=2$, $\bW^{\text{out/in}}_{i\in\{9k\},0}$ being randomly assigned, $D_{\text{in}}=D_{\text{out}}=14$ (vectorized representation of $4\times2$ semi-unitary matrix, Section \ref{vectorize}) and $D_{\text{resv}}=60$.r

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/table.pdf}
\caption{The transmitter obtains feedback for $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$ subcarriers, and uses $\tilde{\bU}_{i,t-1}$, alongwith $f_P,q_P$ \eqref{fpqp} to obtain $\tilde{\bU}_{i,t}$, illustrated in the figure via cyan arrows. For each time instant, the transmitter uses the Cayley method to interpolate $\tilde{\bU}_{i,t}$ at $i\neq\{9k\}$ subcarriers}
\label{table}
\vspace{-8pt}
\end{figure}

\subsection{Quantization error, BER and Achievable Rate Results}
\label{res}{}
Considering mitigation against randomly appearing drastic events in wireless channels, like sudden interferences, and dynamic environment, only 100 channel evolutions are considered for the simulations. Averaging is done over 10 independent channel realizations, to ensure that the simulation results hold in general for the chosen IEEE Pedestrian A channel profile. Hence, the quantization error, BER and achievable rate results are simulated for 1000 channel instances, with 100 channel evolutions of 10 independent channel realizations.

\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{images/qtizErr.pdf}
\caption{Quantization error in terms of chordal distance to the actual precoder $\bU_{\{9k\},t}$ observed at the receiver, from the quantized estimate $\tilde{\bU}_{\{9k\},t}$ at the transmitter, averaged over 8 fed back subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$) and 10 independent channel realizations ($f_DT_s=10^{-4}$)}
\label{qtiz_err}
\end{figure}

From Fig. \ref{qtiz_err} it is evident that the reservoir computing based predictive quantizer is able to reduce the quantization error below what was obtained from the existing temporal correlations based predictive quantizer in \cite{6891198}. In addition, the quantization error is much smoother, which can be explained from the fact that the reservoir computing approach optimizes, $\mathcal{O}_{i,t}$ that captures the long term dynamics of the channel and hence, is able to correct for jittery variations.

>>>>

The output of the reservoir is $\bp_{i,t}$, a vectorized representation of $\bP_{i,t}$.
Therefore, the output of the reservoir is also a $D_{\text{in}}$ dimensional vector ($D_{\text{out}}=D_{\text{in}}$).
The output $\bp_{i,t}$ is obtained using a matrix transformation of the reservoir state $\br_{i,t}$ with the $D_{\text{resv}}\times D_{\text{out}}$ output coupler matrix $\bW^{\text{out}}_{i,t-1}$ viz.
\begin{equation}
\bp_{i,t}=\bW^{\text{out}}_{i,t-1}\br_{i,t}
\label{pred_op}
\end{equation}
\begin{figure*}[ht]
\centering
\includegraphics[width=0.8\textwidth]{images/system2.pdf}
\caption{System Overview: The reservoir computing framework takes the precoder estimate at $t-1$ (viz. $\tilde{\bu}_{i,t-1}$) as input, and executes the forward pass of the predictive quantization framework, shaded in green, to obtain $\tilde{\bu}_{i,t}$. Both the transmitter and receiver utilize the same reservoir (initialized identically), with a minute difference indicated by the cyan shaded block. In the cyan shaded block, receiver finds the appropriate codeword for the $q_P$ function by comparing the prediction with the actual $\bU_{i,t}$, and communicates $\text{fb}_{i,t}$ to allow estimation of $\tilde{\bu}_{i,t}$ at the transmitter. Then, both the receiver and transmitter execute the backward pass (shaded in blue) to optimize $\mathcal{O}_{i,t}$ \eqref{oit}, to ensure identical evolution of both the reservoirs}
\label{res_overview}
\vspace{-10pt}
\end{figure*}
This completes the forward prediction to obtain $\bP_{i,t}$ using only the past observed $\bU_{i,t-1}$ (i.e. $p=1$, Section \ref{section2}). That is,
\begin{equation}
\bP_{i,t}=f_P(\tilde{\bU}_{i,t-1})=\bW^{\text{out}}_{i,t-1}\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})\footnote{For brevity, $\bp_{i,t}$ and $\bP_{i,t}$ are treated to be one and the same}
\label{fwd_pred}
\end{equation}
However, a complete discussion of the prediction algorithm also involves a backward pass, involving the updation of output couplers $\bW^{\text{out}}_{i,t}$, which would follow subsequently.
\subsection{Quantization function $q_P$}
\label{qp}
We now describe the quantization function $q_P(\bP_{i,t},\text{fb}_{i,t})$, which is similar to the one used in \cite{6891198,Gupt1905:Predictive,6545375}.
Central to the quantization algorithm is the fact that tangent spaces local to a point in manifold are linear vector spaces.
Given two points $\bX$ and $\bY$ in $\text{St}(N_T,N_R)$, a lifting operation $\bT^Y_X=\texttt{lift}(\bX,\bY)$, $\texttt{lift}:\text{St}(N_T,N_R)\times \text{St}(N_T,N_R) \to \mathcal{T}_X \text{St}(N_T,N_R)$ yields a tangent from $X$ to $Y$, denoted as $\bT^Y_X \in \mathcal{T}_X \text{St}(N_T,N_R)$, where $\mathcal{T}_X \text{St}(N_T,N_R)$ is the local tangent space at $\bX$.
A corresponding retraction operation $Y=\texttt{retract}(X,\bT^Y_X)$, $\texttt{retract}: \text{St}(N_T,N_R)\times \mathcal{T}_X \text{St}(N_T,N_R) \to \text{St}(N_T,N_R)$ gives back the manifold point obtained by traversing in the tangent direction given by the second argument.
In this work, the chosen lifting-retraction operations are the Cayley exponentials, discussed in \cite{DBLP:journals/corr/abs-1708-00045,Gupt1905:Predictive}. % and used in \cite{Gupt1905:Predictive}.
The Cayley exponential lifting operation maps two points in $\text{St}(N_T,N_R)$, to a $N_T\times N_T$ skew Hermitian matrix that represents the tangent from first point to the other.

The quantization algorithm exploits the linear vector space property of the tangent space and quantizes the local tangent space at the predicted precoder $\bP_{i,t}$, viz $\mathcal{T}_{\bP_{i,t}} \text{St}(N_T,N_R)$.
The codebook for $\mathcal{T}_{\bP_{i,t}} \text{St}(N_T,N_R)$ corresponds to a collection of codewords that represent the different directions (tangents) in $\mathcal{T}_{\bP_{i,t}} \text{St}(N_T,N_R)$.
This codebook is also referred to as the base codebook ($\text{base}^{\mathcal{C}}$) for the quantization function.
The feedback from the receiver indicates the optimum tangent in the $\text{base}^{\mathcal{C}}$, which the transmitter can choose to get closest (in terms of chordal distance metric, $d_s(\bX,\bY)$ for $\bX,\bY \in \text{St}(N_T,N_R)$ detailed in \cite{Gupt1905:Predictive,6891198}) to the actual value $\bU_{i,t}$.
With the optimum direction chosen, the next step is to determine how much to move in that particular chosen direction, viz. the length of the chosen tangent direction.
For this, we adopt the strategy in \cite{6891198}, which controls the magnitude of tangent steps, by having two codebooks $T^{\mathcal{C}}_p,T^{\mathcal{C}}_m$ of different spreads $s_p,s_m$, but same $2^{\mathcal{B}-1}$ base vectors in the $\text{base}^{\mathcal{C}}$, i.e. $T^{\mathcal{C}}_{\{p/m\}}=s_{\{p/m\}}\text{base}^{\mathcal{C}}$ (All codewords $\in \text{base}^{\mathcal{C}}$ multiplied by $s_{\{p/m\}}$ individually).
The two codebooks are concatenated to form a $2^\mathcal{B}$ length codebook $T^\mathcal{C}$.
The receiver finds the optimal index $\text{fb}_{i,t} \in \{1,2,\ldots,2^{\mathcal{B}}\}$ in $T^\mathcal{C}$ by comparing the chordal distance metric ($d_s$), of each codeword to the actual precoder $\bU_{i,t}$ obtained from the SVD of the channel matrix, using \eqref{eqn:cn1}. The receiver then feeds back $\text{fb}_{i,t}$ to the transmitter using $\mathcal{B}$ bits. The transmitter uses the fed back $\text{fb}_{i,t}$ and \eqref{eqn:cn2} to calculate $\tilde{\bU}_{i,t}$,
\label{eqn:cn1}
\text{fb}_{i,t} \gets \text{argmin}_{i \in \{1,2,\ldots,2^{\mathcal{B}}\}}\left (d_s\left (\bU_{i,t},\text{retract}(\bP_{i,t},T^\mathcal{C}[i])\right ) \right)
\tilde{\bU}_{i,t}=q_P(\bP_{i,t},\text{fb}_{i,t})=\text{retract}(\bP_{i,t},T^\mathcal{C}[\text{fb}_{i,t}])
\label{eqn:cn2}
$\text{fb}_{i,t}$ is also used to update the spread of the codebooks $T^{\mathcal{C}}_p,T^{\mathcal{C}}_m$.
Depending on whether $\text{fb}_{i,t} \geq 2^{\mathcal{B}-1}$, i.e. whether the optimum codeword is in $T^{\mathcal{C}}_p$ or $T^{\mathcal{C}}_m$, the scale parameter $s[k]$, which in turn controls values of $s_p, s_m$, is updated accordingly, as

@@

\noindent Having described the technicalities of $f_P,q_P$, we now describe the training process for the reservoir.
Recall that, to obtain $\tilde{\bU}_{i,t}$ from  $\tilde{\bU}_{i,t-1}$, we have

@@

The key idea here is that, as the predicted matrix $\bP_{i,t}$ (the coarse estimate), gets closer to $\tilde{\bU}_{i,t}$ (the refined estimate), the receiver has to quantize even smaller subspaces, and can thus provide a more refined estimate from the feedback $\text{fb}_{i,t}$.
So, for the time instant $t\geq 1$, we train the reservoir output coupler such that the following objective function is optimized:

@@

Here $\lambda>1$ is the history parameter.
We wish to perform the following optimization in order to find the optimum $\bW_{i,t}^{\text{out}}$ from the available estimates till time $t$:

@@

Computing the gradient $\frac{\partial \mathcal{O}_{i,t}}{\partial \bW_{i,t}^{\text{out}}}$ and setting it to the null matrix yields

@@

where $\texttt{outer}(\bx,\by)$ is the outer product between $N_x$ dimensional vector $\bx$ and $N_y$ dimensional vector $\by$ to yield a $N_x\times N_y$ matrix.
Let $\sum_{s=1}^{t}\frac{1}{\lambda^{t-s}}\texttt{outer}(\tilde{\bu}_{i,s},\br_{i,s})$ be $\text{\textbf{num}}_{i,t}$ and $\sum_{s=1}^{t}\frac{1}{\lambda^{t-s}}||\br_{i,s}||^2$ be $\text{den}_{i,t}$, with $\bW_{i,t}^{\text{out}}=\frac{\text{\textbf{num}}_{i,t}}{\text{den}_{i,t}}$.
Observe that

@@

\eqref{wout_upd} makes training of the reservoir easy to implement with low complexity.
We only need to store the matrix $\text{\textbf{num}}_{i,t-1}$, a number $\text{den}_{i,t-1}$. Then, using the current reservoir state $\br_{i,t}$ and $\tilde{\bu}_{i,t}$ obtained from \eqref{fpqp}, $\bW_{i,t}^{\text{out}}$ can be calculated to obtain the predicted coarse estimate of $\tilde{\bu}_{i,t+1}$.
This completes our technical discussion for the reservoir computing based framework for predictive quantization illustrated in Fig. \ref{res_overview}.

@@

Finally, we summarize and highlight the novel aspects of the proposed framework for predictive quantization.
The first key advantage of the proposed approach over the past work \cite{Gupt1905:Predictive,6891198,6545375} is that the prediction function and quantization function are not independent of each other.
To see this, observe \eqref{update_Wt1} to obtain $\bW^{\text{out}}_{i,t-1}$ from $\tilde{\bU}_{i,t-1}=q_P(f_P(\tilde{\bU}_{i,t-2}),\text{fb}_{i,t-1})$, and \eqref{pred} to obtain $\bP_{i,t+1}$
\begin{equation}
\highlight{\bW_{i,t-1}^{\text{out}}}=\frac{\frac{\text{num}_{i,t-2}}{\lambda}+\texttt{outer}(\highlight{q_P}(f_P(\tilde{\bU}_{i,t-2}),\text{fb}_{i,t-1}),\br_{i,t-1})}{\frac{\text{den}_{i,t-2}}{\lambda}+||\br_{i,t-1}||^2}
\label{update_Wt1}
\end{equation}
\begin{equation}
\highlight{\bP_{i,t}}=f_P(\tilde{\bU}_{i,t-1})=\highlight{\bW^{\text{out}}_{i,t-1}}\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\highlight{\tilde{\bu}_{i,t-1}})
\label{pred}
\end{equation}
The dependence between $f_P$ and $q_P$ in our framework is in the sense that the backward pass ensures that prediction function `learns' the correlations brought about by $q_P$, in addition to temporal correlations. %Better framing possible
This is done by optimizing $\bW_{i,t}^{\text{out}}$ such that the prediction output to be as close to the output from $q_P$ (Objective function $\mathcal{O}_{i,t}$).
The updated $\bW_{i,t}^{\text{out}}$ is then used in the forward pass to predict $\bP_{i,t+1}$, which is likely closer to $\tilde{\bU}_{i,t+1}$ than $\bP_{i,t}$ was to $\tilde{\bU}_{i,t}$ due to optimization of $\mathcal{O}_{i,t}$.

\textcolor{blue}{Another advantage of the proposed scheme is the ease of training the reservoir, captured in \eqref{res_upd}.
By storing just one matrix $\text{\textbf{num}}_{i,t}$ and a number $\text{den}_{i,t}$, we capture the entire history of the precoding matrices, with past values weighted by $\lambda$.
This is a departure from previous schemes \cite{Gupt1905:Predictive,6891198,6545375} which store the past $n$ precoders precoders in a $n$ sized cyclic buffer ($n$ also has to be pre-decided).
Reservoir computing, thus proposes an easier to train (data centric) scheme, which is not typical in data centric ML based schemes.}

\section{Simulation Results}
\label{section4}
\subsection{Simulation setting considered}
\label{setting}
\noindent We simulate both our predictive quantization algorithm and the one presented in \cite{6891198} for the IEEE Pedestrian-A channel, with $N_T=4$, $N_R=2$. 
The channel matrices $\bH_{i,t}$ are generated using Jake's model for realistic simulations, from the IT++ library through the python wrapper py-itpp \cite{ViditPy}. 
Although in both our work and \cite{6891198}, only temporal correlations are exploited, we compare the BER, quantization error and logsum rate results for the complete MIMO-OFDM setting with $N=64$ of subcarriers, and channels are fed back for $8$ evenly spaced subcarriers, indexed starting at 0, viz. $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$).
The precoding matrices at subcarriers were channels are not fed back ($i\neq\{9k\}$) are estimated via the Cayley Interpolation method presented in \cite{Gupt1905:Predictive}.
Feedback is available for $i=\{9k\}$ subcarriers (i.e. transmitter obtains $\text{fb}_{\{9k\},t}$) for predictive quantization.
The setting is visually illustrated in Fig. \ref{table}.
A 6 bit codebook generated using the Lloyd codebook algorithm for the Stiefel Manifold in \cite{6678348} is used to feedback $\tilde{\bU}_{\{9k\},0}$ and differential quantization (i.e. $\bP_{i,t}=\tilde{\bU}_{i,t-1}$) is used for 10 time instances to provide provide the initial training data for the reservoir and for overcoming initial transient response of the reservoir, as discussed in \cite{mosleh2017brain}.
A 6 bit base codebook that is used for the quantization function $q_P$ is generated according to the method presented in Section \ref{qp} and \cite{Gupt1905:Predictive}.
Hence, for both initialization and subsequent feedback, $\text{fb}_{i,t}$ can be encoded using 6 bits (i.e. $\mathcal{B}=6$ Section \ref{section2}).
The bit budget per OFDM frame is thus 48 bits, for total 8 equally spaced fed back subcarriers $\{9k, k \in \{0,1,\ldots,7\}\}$.

For predicting channels at each of the $8$ fed-back subcarriers, we consider $8$ separate reservoirs, each of them evolved separately, with $i\in\{9k, k \in \{0,1,\ldots,7\}\}$ in both the forward pass prediction and quantization ($f_P,q_P$) \eqref{fpqp} and backward pass to update $\bW_{i,t}^{\text{out}}$ \eqref{backpass}. Unlike \cite{Gupt1905:Predictive}, we do not exploit frequency correlations for predictive quantization, and the $8$ reservoirs for prediction of $8$ fed-back subcarriers are kept independent of each other.
All these 8 reservoirs are initialized with $\bA_{i\in\{9k\}}$ being an Erdos-Reny\"i graph, with probability of edge connections being $0.2$, $\lambda=2$, $\bW^{\text{out/in}}_{i\in\{9k\},0}$ being randomly assigned, $D_{\text{in}}=D_{\text{out}}=14$ (vectorized representation of $4\times2$ semi-unitary matrix, Section \ref{vectorize}) and $D_{\text{resv}}=60$.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/table.pdf}
\caption{The transmitter obtains feedback for $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$ subcarriers, and uses $\tilde{\bU}_{i,t-1}$, along with $f_P,q_P$ \eqref{fpqp} to obtain $\tilde{\bU}_{i,t}$, illustrated in the figure via cyan arrows. For each time instant, the transmitter uses the Cayley method to interpolate $\tilde{\bU}_{i,t}$ at $i\neq\{9k\}$ subcarriers.}
\label{table}
\vspace{-8pt}
\end{figure}

\subsection{Quantization error, BER and Achievable Rate Results}
\label{res}{}
To mitigate errors in tracking in the presence of randomly appearing drastic events in wireless channels, such as sudden interferences and dynamic environments, only 100 channel evolutions are considered for the simulations.
Averaging is performed over 10 independent channel realizations, to ensure that the simulation results hold in general for the chosen IEEE Pedestrian A channel profile. Therefore, the quantization error, BER and achievable rate results are averaged over 1000 channel instances, with 100 channel evolutions of 10 independent channel realizations.
\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{images/qtizErr.pdf}
\caption{Quantization error in terms of chordal distance to the actual precoder $\bU_{\{9k\},t}$ observed at the receiver, from the quantized estimate $\tilde{\bU}_{\{9k\},t}$ at the transmitter, averaged over 8 fed back subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$) and 10 independent channel realizations (normalized Doppler $f_DT_s=10^{-4}$).}
\label{qtiz_err}
\end{figure}
From Fig. \ref{qtiz_err} it is evident that the reservoir computing based predictive quantizer is able to reduce the quantization error to below what was obtained from the existing temporal correlation based predictive quantizer in \cite{6891198}. In addition, the quantization error is much smoother, which can be explained from the fact that the reservoir computing approach optimizes, $\mathcal{O}_{i,t}$ that captures the long term dynamics of the channel and hence, is able to correct for jittery variations.

We simulate the (uncoded QPSK) BER performance for precoders obtained after interpolation ($\tilde{\bU}_{i\neq\{9k\},t}$) and quantization ($\tilde{\bU}_{\{9k\},t}$), for all the 63 subcarriers.
Observe from Fig. \ref{BER} that reservoir computing framework is able to achieve substantial improvements in $\text{E}_\text{b}\text{N}_0$ levels at $\text{BER}\leq 10^{-3}$.
Particularly, for $\text{BER}=10^{-4}$, we observe around $5$ dB improvement.
In addition, substantial improvement is obtained with both the predictive quantization schemes over the independent quantization scheme with Lloyd codebook for precoders \cite{6678348}, using the same number of bits.
